---
title: "P8105 Homework 3"
author: Heidi Lumish
output: github_document
---

```{r setup, include=FALSE}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1.0,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

First we will load in the Instacart data from the p8105.datasets library.

```{r}
data("instacart")
```

#### Summary of the Instacart dataset

Next, we will summarize the dataset.

The Instacart dataset provides detailed information on grocery orders in 2017. The dataset has `r ncol(instacart)` variables, and it describes `r nrow(instacart)` products from `r length(unique(pull(instacart, user_id)))` unique users. `r round(sum(pull(instacart, reordered))/nrow(instacart)*100, digits = 2)`% products had been ordered by the user in the past. The median time since the prior order for each customer was `r median(pull(instacart, "days_since_prior_order"))` (IQR `r quantile(pull(instacart, "days_since_prior_order"), 0.25)`, `r quantile(pull(instacart, "days_since_prior_order"), 0.75)`) days. There were `r length(unique(pull(instacart, product_name)))` unique products ordered from `r length(unique(pull(instacart, aisle_id)))` different aisles.

### Specifics of the Instacart dataset

#### How many aisles are there, and which aisles are the most items ordered from?

There are `r length(unique(pull(instacart, aisle_id)))` aisles, and the most items are ordered from aisle. Using the following code chunk, we can determine that the most items are ordered from the aisle fresh vegetables, fresh fruits, packaged vegetables, yogurt, and packaged cheese.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    aisle_count = n()
  ) %>% 
  slice_max(aisle_count, n = 5)
```

#### Plot showing the number of items ordered in each aisle

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    aisle_count = n()
  ) %>% 
  filter(aisle_count >10000) %>% 
  ggplot(aes(x = reorder(aisle, aisle_count), y = aisle_count)) +
  geom_col(width = 0.7) +
  coord_flip() + 
  labs(
    title = "Number of products ordered per aisle",
    x = "Aisle name",
    y = "Number of products")
```

#### Table of three most popular items
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(
    aisle == "baking ingredients" |
      aisle == "dog food care" |
      aisle == "packaged vegetables fruits") %>%
  group_by(product_name) %>% 
  mutate(
    product_count = n()) %>%
  select(aisle, product_name, product_count) %>% 
  distinct() %>% 
  group_by(aisle) %>%
  slice_max(product_count, n = 3) %>% 
  mutate(
    Rank = order(product_count, decreasing = TRUE)
  ) %>%
  unite("product", product_name, product_count, sep = ": ") %>% 
  mutate(
    aisle = str_to_title(aisle)
  ) %>% 
  pivot_wider(
    names_from = "aisle",
    values_from = "product"
    ) %>% 
  knitr::kable(format = "simple",
               caption = "Top three most popular items by aisle")
```

#### Pink Lady Apples and Coffee Ice Cream

The following is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart %>% 
  filter(
    product_name == "Pink Lady Apples" |
    product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hour = round(mean(order_hour_of_day), digits = 1)
  ) %>% 
  mutate(Day = case_when(
    order_dow == 0 ~ "Sunday",
    order_dow == 1 ~ "Monday",
    order_dow == 2 ~ "Tuesday",
    order_dow == 3 ~ "Wednesday",
    order_dow == 4 ~ "Thursday",
    order_dow == 5 ~ "Friday",
    order_dow == 6 ~ "Saturday")
  ) %>% 
  pivot_wider(
    names_from = "product_name",
    values_from = "mean_hour"
  ) %>% 
  select(-order_dow) %>% 
    knitr::kable(format = "simple",
                 caption = "Mean hour for ordering products by day of the week")
```

## Problem 2

#### Load and clean the BRFSS dataset

First we will load the BRFSS dataset from the p8105.datasets package.

```{r}
data("brfss_smart2010")
```

Next we will clean the data.

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  mutate(
    state = locationabbr,
    county = locationdesc
  ) %>% 
  select(-locationabbr, -locationdesc) %>% 
  relocate(year, state, county) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent")) %>%
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))
  )
```

#### BRFSS dataset questions

In 2002, which states were observed at 7 or more locations? What about in 2010?

```{r}
brfss %>% 
  filter(year == "2002") %>%
  select(state, county) %>% 
  distinct() %>% 
  group_by(state) %>% 
  summarize(count = n()) %>% 
  filter(count >= 7)

brfss %>% 
  filter(year == "2010") %>%
  select(state, county) %>% 
  distinct() %>% 
  group_by(state) %>% 
  summarize(count = n()) %>% 
  filter(count >= 7)
```

Construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state (that is, make a plot showing a line for each state across years – the geom_line geometry and group aesthetic will help).

```{r}
excellent = brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  mutate(
    mean_data_value = mean(data_value)
  ) %>% 
  select(year, state, mean_data_value)

excellent %>% 
  ggplot(aes(x = year, y = mean_data_value, color = state)) +
  geom_point(size = .2) + 
  geom_line(aes(group = state), alpha = .3)
```

Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>% 
  filter(
    year == c("2006", "2010"),
    state == "NY") %>% 
  ggplot(aes(x = response, y = data_value, color = county)) +
  geom_point() + 
  facet_grid(.~year)
```

## Problem 3
This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The data can be downloaded here. In this spreadsheet, variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes.

First we will load in and tidy the accelerometer data.

```{r}
accel = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    day_type = case_when(
      day == "Sunday" ~ "weekend",
      day == "Monday" ~ "weekday",
      day == "Tuesday" ~ "weekday",
      day == "Wednesday" ~ "weekday",
      day == "Thursday" ~ "weekday",
      day == "Friday" ~ "weekday",
      day == "Saturday" ~ "weekend")
  ) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    names_prefix = "activity_",
    values_to = "counts"
  )
```

Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel %>% 
  group_by(day_id) %>% 
  mutate(total_activity = sum(counts)) %>% 
  ungroup() %>% 
  select(week, day, total_activity) %>% 
  distinct() %>% 
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  )
```

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel %>% 
  ggplot(aes(x = day_id, y = counts, color = day, shape = day_type)) + 
  geom_point()
```


