---
title: "P8105 Homework 3"
author: Heidi Lumish
output: github_document
---

```{r setup, include=FALSE}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1.0,
  out.width = "100%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

First we will load in the Instacart data from the p8105.datasets library.

```{r}
data("instacart")
```

#### Summary of the Instacart dataset

Next, we will summarize the dataset.

The Instacart dataset provides detailed information on grocery orders in 2017. The dataset has `r ncol(instacart)` variables, and it describes `r nrow(instacart)` products from `r length(unique(pull(instacart, user_id)))` unique users. `r round(sum(pull(instacart, reordered))/nrow(instacart)*100, digits = 2)`% products had been ordered by the user in the past. The median time since the prior order for each customer was `r median(pull(instacart, "days_since_prior_order"))` (IQR `r quantile(pull(instacart, "days_since_prior_order"), 0.25)`, `r quantile(pull(instacart, "days_since_prior_order"), 0.75)`) days. There were `r length(unique(pull(instacart, product_name)))` unique products ordered from `r length(unique(pull(instacart, aisle_id)))` different aisles.

### Specifics of the Instacart dataset

#### How many aisles are there, and which aisles are the most items ordered from?

There are `r length(unique(pull(instacart, aisle_id)))` aisles, and the most items are ordered from aisle. Using the following code chunk, we can determine that the most items are ordered from the aisle fresh vegetables, fresh fruits, packaged vegetables, yogurt, and packaged cheese.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    aisle_count = n()
  ) %>% 
  slice_max(aisle_count, n = 5)
```

#### Plot showing the number of items ordered in each aisle

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    aisle_count = n()
  ) %>% 
  filter(aisle_count >10000) %>% 
  ggplot(aes(x = reorder(aisle, aisle_count), y = aisle_count)) +
  geom_col(width = 0.7) +
  coord_flip() + 
  labs(
    title = "Number of products ordered per aisle",
    x = "Aisle name",
    y = "Number of products")
```

#### Table of three most popular items
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(
    aisle == "baking ingredients" |
      aisle == "dog food care" |
      aisle == "packaged vegetables fruits") %>%
  group_by(product_name) %>% 
  mutate(
    product_count = n()) %>%
  select(aisle, product_name, product_count) %>% 
  distinct() %>% 
  group_by(aisle) %>%
  slice_max(product_count, n = 3) %>% 
  mutate(
    Rank = order(product_count, decreasing = TRUE)
  ) %>%
  unite("product", product_name, product_count, sep = ": ") %>% 
  mutate(
    aisle = str_to_title(aisle)
  ) %>% 
  pivot_wider(
    names_from = "aisle",
    values_from = "product"
    ) %>% 
  knitr::kable(format = "simple",
               caption = "Top three most popular items by aisle")
```

#### Pink Lady Apples and Coffee Ice Cream

The following is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart %>% 
  filter(
    product_name == "Pink Lady Apples" |
    product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hour = round(mean(order_hour_of_day), digits = 1)
  ) %>% 
  mutate(Day = case_when(
    order_dow == 0 ~ "Sunday",
    order_dow == 1 ~ "Monday",
    order_dow == 2 ~ "Tuesday",
    order_dow == 3 ~ "Wednesday",
    order_dow == 4 ~ "Thursday",
    order_dow == 5 ~ "Friday",
    order_dow == 6 ~ "Saturday")
  ) %>% 
  pivot_wider(
    names_from = "product_name",
    values_from = "mean_hour"
  ) %>% 
  select(-order_dow) %>% 
    knitr::kable(format = "simple",
                 caption = "Mean hour for ordering products by day of the week")
```

## Problem 2

#### Load and clean the BRFSS dataset

First we will load the BRFSS dataset from the p8105.datasets package.

```{r}
data("brfss_smart2010")
```

Next we will clean the data.

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  filter(topic == "Overall Health") %>% 
  filter(response %in% c("Poor", "Fair", "Good", "Very Good", "Excellent")) %>%
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))
  )
```





