---
title: "P8105 Homework 3"
author: Heidi Lumish
output: github_document
---

```{r setup, include=FALSE}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1.0,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

First we will load in the Instacart data from the p8105.datasets library.

```{r}
data("instacart")
```

#### Summary of the Instacart dataset

Next, we will summarize the dataset.

The Instacart dataset provides detailed information on grocery orders in 2017. The dataset has `r ncol(instacart)` variables, and it describes `r nrow(instacart)` products from `r length(unique(pull(instacart, user_id)))` unique users. `r round(sum(pull(instacart, reordered))/nrow(instacart)*100, digits = 2)`% products had been ordered by the user in the past. The median time since the prior order for each customer was `r median(pull(instacart, "days_since_prior_order"))` (IQR `r quantile(pull(instacart, "days_since_prior_order"), 0.25)`, `r quantile(pull(instacart, "days_since_prior_order"), 0.75)`) days. There were `r length(unique(pull(instacart, product_name)))` unique products ordered from `r length(unique(pull(instacart, aisle_id)))` different aisles.

### Specifics of the Instacart dataset

#### How many aisles are there, and which aisles are the most items ordered from?

There are `r length(unique(pull(instacart, aisle_id)))` aisles, and the most items are ordered from aisle. Using the following code chunk, we can determine that the most items are ordered from the aisle fresh vegetables, fresh fruits, packaged vegetables, yogurt, and packaged cheese.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    aisle_count = n()
  ) %>% 
  slice_max(aisle_count, n = 5)
```

#### Plot showing the number of items ordered in each aisle

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    aisle_count = n()
  ) %>% 
  filter(aisle_count >10000) %>% 
  ggplot(aes(x = reorder(aisle, aisle_count), y = aisle_count)) +
  geom_col(width = 0.7) +
  coord_flip() + 
  labs(
    title = "Number of products ordered per aisle",
    x = "Aisle name",
    y = "Number of products")
```

#### Table of three most popular items
Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(
    aisle == "baking ingredients" |
      aisle == "dog food care" |
      aisle == "packaged vegetables fruits") %>%
  group_by(product_name) %>% 
  mutate(
    product_count = n()) %>%
  select(aisle, product_name, product_count) %>% 
  distinct() %>% 
  group_by(aisle) %>%
  slice_max(product_count, n = 3) %>% 
  mutate(
    Rank = order(product_count, decreasing = TRUE)
  ) %>%
  unite("product", product_name, product_count, sep = ": ") %>% 
  mutate(
    aisle = str_to_title(aisle)
  ) %>% 
  pivot_wider(
    names_from = "aisle",
    values_from = "product"
    ) %>% 
  knitr::kable(format = "simple",
               caption = "Top three most popular items by aisle")
```

#### Pink Lady Apples and Coffee Ice Cream

The following is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.

```{r}
instacart %>% 
  filter(
    product_name == "Pink Lady Apples" |
    product_name == "Coffee Ice Cream") %>% 
  group_by(product_name, order_dow) %>% 
  summarize(
    mean_hour = round(mean(order_hour_of_day), digits = 1)
  ) %>% 
  mutate(Day = case_when(
    order_dow == 0 ~ "Sunday",
    order_dow == 1 ~ "Monday",
    order_dow == 2 ~ "Tuesday",
    order_dow == 3 ~ "Wednesday",
    order_dow == 4 ~ "Thursday",
    order_dow == 5 ~ "Friday",
    order_dow == 6 ~ "Saturday")
  ) %>% 
  pivot_wider(
    names_from = "product_name",
    values_from = "mean_hour"
  ) %>% 
  select(-order_dow) %>% 
    knitr::kable(format = "simple",
                 caption = "Mean hour for ordering products by day of the week")
```

## Problem 2

#### Load and clean the BRFSS dataset

First we will load the BRFSS dataset from the p8105.datasets package.

```{r}
data("brfss_smart2010")
```

Next we will clean the data.

```{r}
brfss = brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  mutate(
    state = locationabbr,
    county = locationdesc
  ) %>% 
  select(-locationabbr, -locationdesc) %>% 
  relocate(year, state, county) %>% 
  filter(topic == "Overall Health") %>% 
  filter(response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>%
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))
  )
```

#### BRFSS dataset questions

###### Observed at 7 or more locations
In 2002, which states were observed at 7 or more locations? What about in 2010?

Based on the code chunk below, we can see that the there were 6 states (CT, FL, MA, NC, NJ, and PA) that were observed at 7 or more locations in 2002. By comparison, in 2010, there 14 states (CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, and WA) that were observed at 7 or more locations.

```{r}
brfss %>% 
  filter(year == "2002") %>%
  select(state, county) %>% 
  distinct() %>% 
  group_by(state) %>% 
  summarize(count = n()) %>% 
  filter(count >= 7)

brfss %>% 
  filter(year == "2010") %>%
  select(state, county) %>% 
  distinct() %>% 
  group_by(state) %>% 
  summarize(count = n()) %>% 
  filter(count >= 7)
```

###### Excellent responses
Construct a dataset that is limited to *Excellent* responses, and contains year, state, and a variable that averages the data_value across locations within a state. Make a “spaghetti” plot of this average value over time within a state.

```{r}
excellent = brfss %>% 
  filter(response == "Excellent") %>% 
  group_by(state, year) %>% 
  mutate(
    mean_data_value = mean(data_value)
  ) %>% 
  select(year, state, mean_data_value)

excellent %>% 
  ggplot(aes(x = year, y = mean_data_value, color = state)) +
  geom_point(size = .2) + 
  geom_line(aes(group = state), alpha = .3)
```

###### Distribution of responses in 2006 and 2010 in NY state
Make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.

```{r}
brfss %>% 
  filter(
    year == c("2006", "2010"),
    state == "NY") %>% 
  ggplot(aes(x = data_value, fill = response)) +
  geom_density() + 
  facet_grid(.~year) +
  labs(
    title = "Distribution of data values by response in NY",
    x = "Data value",
    y = "Density")
```

This density plot shows that a response of "Poor" correlated with the lowest data values, both in 2006 and 2010. A response of "Fair" similarly correlated with lower data values, though not quite as low as for "Poor." In 2006, there was overlap between the data values for "Good," "Very good," and "Excellent responses. In 2010, the highest data values coresponsed with "Good" and "Very good" responses.

## Problem 3
This problem uses five weeks of accelerometer data collected on a 63 year-old male with BMI 25, who was admitted to the Advanced Cardiac Care Center of Columbia University Medical Center and diagnosed with congestive heart failure (CHF). The variables activity.* are the activity counts for each minute of a 24-hour day starting at midnight.

#### Load and tidy the accelerometer data
First we will load, tidy, and wrangle the accelerometer data.

```{r}
accel = read_csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    day_type = case_when(
      day == "Sunday" ~ "weekend",
      day == "Monday" ~ "weekday",
      day == "Tuesday" ~ "weekday",
      day == "Wednesday" ~ "weekday",
      day == "Thursday" ~ "weekday",
      day == "Friday" ~ "weekday",
      day == "Saturday" ~ "weekend"),
    week = as.numeric(as.character(week))
  ) %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    names_prefix = "activity_",
    values_to = "counts"
  ) %>% 
  mutate(counts = as.numeric(counts))
```

#### Summary of the dataset
Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

The tidied accelerometer dataset has `r ncol(accel)` variables and `r nrow(accel)` observations. It includes data from week `r min(pull(accel, week))` to week `r max(pull(accel, week))` of the study. The dataset includes activity counts for each minute of 24-hour day. The mean activity count per minute during the study period was `r mean(pull(accel, counts))`. 

#### Total activity per day
Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel %>% 
  group_by(day_id) %>% 
  mutate(total_activity = sum(counts)) %>% 
  ungroup() %>% 
  select(week, day, total_activity) %>% 
  distinct() %>% 
  pivot_wider(
    names_from = "day",
    values_from = "total_activity"
  )
```

#### 24-hour activity pattern by day
Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel %>% 
  mutate(
    activity = as.numeric(as.character(activity)),
    activity = activity/60
    ) %>% 
  ggplot(aes(x = activity, y = counts, color = day)) + 
  geom_point() +
  labs(
    title = "24-hour activity pattern by day of the week",
    x = "Hour",
    y = "Activity")
```

Based on this plot, we can see that there are two peaks of activity during the day, around 10 am and 8 pm. The most activity tended to occur on the weekends and on Fridays.

