---
title: "P8105 Homework 3"
author: Heidi Lumish
output: github_document
---

```{r setup, include=FALSE}
library(p8105.datasets)
library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

First we will load in the Instacart data from the p8105.datasets library.

```{r}
data("instacart")
```

#### Summary of the Instacart dataset

Next, we will summarize the dataset.

The Instacart dataset provides detailed information on grocery orders in 2017. The dataset has `r ncol(instacart)` variables, and it describes `r nrow(instacart)` products from `r length(unique(pull(instacart, user_id)))` unique users. `r round(sum(pull(instacart, reordered))/nrow(instacart)*100, digits = 2)`% products had been ordered by the user in the past. The median time since the prior order for each customer was `r median(pull(instacart, "days_since_prior_order"))` (IQR `r quantile(pull(instacart, "days_since_prior_order"), 0.25)`, `r quantile(pull(instacart, "days_since_prior_order"), 0.75)`) days. There were `r length(unique(pull(instacart, product_name)))` unique products ordered from `r length(unique(pull(instacart, aisle_id)))` different aisles.

#### Specifics of the Instacart dataset

How many aisles are there, and which aisles are the most items ordered from?

There are `r length(unique(pull(instacart, aisle_id)))` aisles, and the most items are ordered from aisle. Using the following code chunk, we can determine that the most items are ordered from the aisle fresh vegetables, fresh fruits, packaged vegetables, yogurt, and packaged cheese.

```{r}
instacart %>% 
  group_by(aisle) %>% 
  summarize(
    aisle_count = n()
  ) %>% 
  slice_max(aisle_count, n = 5)
```


